---
title: "Homework 2"
author: "Simon G. Brauer"
date: "Sunday, November 08, 2015"
output: pdf_document
---

##Instructions

1. Take a data set that you are currently working on, and select a small number of variables of interest from it-fewer than ten. You can do this manually if you like-e.g. in Stata, or Excel-or you can use the subset() function to select data from the full dataset. Either way is fine.
2. Using RStudio, create either an RMarkdown file (preferred) or an .R file (also OK) that does the following:
- Get the data into R so you can work with it
- Produce numerical summaries of the variables with summary() or similar
- Use ggplot to produce histograms or some other appropriate 1-D distributional summary of each variable separately
- Use ggplot and GGally packages together to produce a pairs or generalized pairs plot of all the variables together
- Select two or three variables and look at their bivariate relationships more closely. E.g. you might produce a scatterplot & smoother of two continuous variables, or see how the distribution of some continuous variable (e.g. income) varies by a categorical variable (e.g. level of education) 
3. Produce a PDF of the results and send it to me, OR use RStudio's publication feature to make a publicly accessible HTML version on its RPubs service. If you use RMarkdown this should be straightforward. If you use an .R script, save the graphical output into individual files and include them in a single PDF document. Your document should contain brief linking information (i.e. literally one or two sentences) explaining what each variable and plot is. You can do as many plots as you like, but at a minimum please show me summaries of each variable individually, a pairs plot, and at least two bivariate or multivariate plots.

```{r, message = FALSE, warning = FALSE}
setwd("C:/Users/Simon/OneDrive/Documents/Github/Data-Visualization/")
library(foreign)
library(ggplot2)
library(GGally)
NCS <- read.dta("NCS.dta")
attach(NCS)
```

I am using data from the National Congregations Study (NCS). I've included variables measuring religious tradition, size of the congregation, clergy's race, congregation's income, the percentage of homosexual attendees, restrictions on homosexual membership and leadership, and how liberal or conservative the congregation is.

```{r, warning = FALSE}
head(NCS)
```

##Univariate statistics

Note that all results are unweighted. Because the NCS is a hypernetwork sample, the distribution is representative of *attendees* and not congregations.

For example, the average attendee has 776 adults in their congregation (`numadlts`). Likewise, roughly half (1,388 of `r 1388 + 1170 + 258`) of attendees in the US are in congregations in which homosexuals are allowed to become members (`mbrgay`)

```{r, echo = FALSE, message = FALSE, warning = FALSE}
###First, I graph the discrete variables with histograms
discrete.description <- c("Catholic parishes make up the majority of the NCS data, followed by Baptists. No other group has more than 500 congregations represented.",
                          "Clergy are overwhelmingly white in the NCS.",
                          "Nearly half of individuals are in a congregation that allows homosexual members.",
                          "Most individuals are in a congregation that restricts homosexuals from becoming leaders.",
                          "Most individuals are in congregations that are conservative or moderate. Relatively few are in liberal congregations.")

discrete.title <- c("Congregation's Religious Tradition", "Clergy's Race", "Congregation Allows Homosexual Members", "Congregation Allows Homosexual Leaders", "Congregation's Political Position")
```

```{r, warning = FALSE}
counter <- 1
angle1 <- c(15, 15, 0, 0, 15)
horz <- c(1, 1, 0.5, 0.5, 1)
for(i in list(DENCODE3, clerrace, mbrgay, ldrgay, libcon)){
  print(discrete.title[counter], quote = FALSE)
  print(discrete.description[counter], quote = FALSE)
  print(summary(i), quote = FALSE)
  
  temp.fig <- ggplot(data = NCS) + 
    geom_histogram(aes(x = i)) +
    ggtitle(discrete.title[counter]) + 
    ylab("") +
    xlab("") +
    theme(axis.text.x = element_text(angle = angle1[counter], hjust = horz[counter]))
  print(temp.fig)
  counter <- counter + 1
}
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
###Next, I graph the continuous variables with density plots
continuous.description <- c("The average atendee has 776 adults that regularly attend their congregation. The median attendee has 250, indicating that the distribution is highly right-skewed.",
                            "The average attendee is in a congregation that brings in just under $1,000,000 a year, though the median attendee is in a congregation that brings in $300,000",
                            "While the average attendee's congregation is 1.4% made up of homosexual members, the median is 0, indicating that most attendees have no homosexual members in their congregation, or at least that the respondent did not know of or chose not to report any.")

continuous.title <- c("Logged number of regularly attending adults", "Logged total income", "Logged percentage attendees that are homosexual")

continuous.label <- c("ln(adults)", "ln(dollars)", "ln(percentage)")
```

```{r, warning = FALSE}
counter <- 1
for(i in list(numadlts, income, gaypct)){
  print(continuous.title[counter], quote = FALSE)
  print(continuous.description[counter], quote = FALSE)
  print(summary(i), quote = FALSE)
  
  temp.fig <- ggplot(data = NCS) + 
    geom_density(aes(x = log(i))) +
    ggtitle(continuous.title[counter]) + 
    ylab("") +
    xlab(continuous.label[counter])
  print(temp.fig)
  counter <- counter + 1
}
```

##Bivariate relationships

First, I produce a pairs plot of all of my variables to look for concerning patterns. 

```{r, warning = FALSE}
NCS.log <- NCS
NCS.log$numadlts <- log(NCS$numadlts)
NCS.log$income <- log(NCS$income)
NCS.log$gaypct <- log(NCS$gaypct)

ggpairs(NCS.log, 
        upper = list(continuous = "smooth", combo = "box", discrete = "facetbar"),
        lower = list(continuous = "blank", combo = "blank", discrete = "blank"))
```

While not seemingly significantly different, it looks like there's some variation in the percentage of homosexuals in congregations by religious tradition. I look at this further, first, with a density plot to show the distributions, and second, with box plots.

```{r, echo = FALSE, warning = FALSE}
ggplot(data = NCS.log) + 
  geom_density(aes(x = gaypct)) + 
  facet_wrap(~DENCODE3) +
  ggtitle("Logged percentage homosexual attendees across eleven religious traditions") +
  ylab("") +
  xlab("Ln(percent homosexual attendees)")

ggplot(data = NCS.log) + 
  geom_boxplot(aes(x = DENCODE3, y = gaypct)) + 
  coord_flip() +
  ggtitle("Logged percentage homosexual attendees across eleven religious traditions") +
  ylab("ln(percent homosexual attendees)") +
  xlab("")
```

