stall.---
title: "Notes - Class 1"
author: "Simon G. Brauer"
date: "Tuesday, October 06, 2015"
output: pdf_document
---
#Class 1: October 6th, 2015
##Websites for the class
http://socviz.github.io/soc880/

http://github.com/socviz/soc880

##First sample code (sort of)
```{r}
data <- c(1, 1, 4, 1, 1, 4, 1)
data

x <- rnorm(n = 100)
y <- x + rnorm(n = 100)
```

```{r, echo = FALSE}
library(ggplot2)
testmod <- lm(x ~ y)
ggplot() +
  geom_point(aes(x = x, y = y)) +
  geom_abline(intercept = 0.057, slope = 0.509)
```

##Details on R
R has a working directory
```{r}
getwd()
```

How Kieran organizes his working directory...

1. data folder
2. doc folder
3. figures folder

* r file in root

everything in R is an object...
```{r}
(my_numbers <- c(1, 2, 3, 1, 3,5, 25))
summary(my_numbers)
```

every object has a class
```{r}
class(my_numbers)
#even functions have classes
class(summary)
```

Objects can hold the output of functions
```{r}
(my_summary <- summary(my_numbers))
class(my_summary)
```

Examples in R
```{r}
my_numbers * 2
table(my_numbers)
sd(my_numbers)
```

you can get information in different ways
```{r}
str(my_numbers)
str(my_summary)
str(summary)
```

#why are charts sometimes bad?

*Pure aesthetics: presentation is fine but ugly

*Figure shows you less than it should or could

*Figure shows more than it should

*Figure can be misleading or misrepresent the data, accidentally or purposefully

Consider the actual amount of information you're presenting and conside simplifying to emphasizeit (instead the background, coloring, etc.)

*Look at William Cleveland's work on presenting data*

His thoughts:

* It's difficult to compare without common baselines
* People misjudge areas like crazy
* Acute angles are underestimated, obtuse are overestimated
* You can't distinguish data points if they are overlapping
* Colors are good for distinguishing categories but not for continuous variables
* Difficult at interpreting curves because our perception is influenced by horizontal space

###Let's get some data...
```{r}
library(devtools)
gapminder.url <- "https://raw.githubusercontent.com/socviz/soc880/master/data/gapminder.csv"

## What type of object is this?
class(gapminder.url)
data <- read.csv(url(gapminder.url))
class(data)
str(data)
head(data)
```

let's plot it!
```{r}
p <- ggplot(data = data, aes(x = gdpPercap, y = lifeExp))
lm(data = data, formula = lifeExp ~ gdpPercap)
p + geom_point() + geom_abline(intercept = 54, slope = 0.0007)
```

##Homework for next time
1. Create an RMarkdown file for your work
2. Look again at the data
3. Put lifeExp on the x-axis and gdpPercap on the y-axis
4. plot pop on the x-axis and gdpPercap on the y-axis
5. Plot year on the x-axis and any continuous variable on the y-axis

#Homework
```{r}
##Put lifeEXP on the x-axis and gdpPercap on the y-axis
ggplot(data = data, aes(x = lifeExp, y = gdpPercap)) + geom_point() + labs(title = "GDP Per Capita plotted Across Life Expectancy")

##Put pop on the x-axis and gdpPercap on the y-axis
ggplot(data = data, aes(x = pop, y = gdpPercap)) + geom_point() + labs(title = "GDP Per Capita Plotted Across Population")

##Plot year on the x-axis and any continuous variable on the y-axis
ggplot(data = data, aes(x = year, y = pop)) + geom_point() + labs(title = "Population Plotted Across Years")
```
