---
title: "Class 3 - 10/27/2015"
author: "Simon G. Brauer"
date: "Tuesday, October 27, 2015"
output: html_document
---

```{r include = FALSE}
require(ggplot2)

gapminder.url <- "https://raw.githubusercontent.com/socviz/soc880/master/data/gapminder.csv"
my.data <- read.csv(url(gapminder.url))
dim(my.data)

head(my.data)
```

This plot is wrong...

```{r}
p <- ggplot(my.data, aes(x = gdpPercap, y = lifeExp))
p + geom_line() + scale_x_log10()
```

Is this better?

```{r}
p <- ggplot(my.data, aes(x = year, y = lifeExp))
p + geom_line() + 
    scale_x_log10()
```

OK, one more time.

```{r}
p <- ggplot(my.data, aes(x = year, y = lifeExp))
p + geom_line(aes(group = country))
```

Ooooooone more time...

```{r}
p <- ggplot(my.data, aes(x = year, y = lifeExp))
p + geom_line(aes(group = country, color = continent))
```

Let's just look at Asia.

```{r}
p <- ggplot(my.data, aes(x = year, y = lifeExp))
p + geom_line(aes(group = country, color = continent=="Asia"))
#Shoutout from Kieran. Yesssssss.
```

Hmm, interesting. Let's make it more attractive.

```{r}
p <- ggplot(my.data, aes(x = year, y = lifeExp))
p + geom_line(aes(group = country, color = continent=="Asia")) +
    scale_color_manual(values = c("gray70", "firebrick"), 
                       labels = c("All other countries", "Asian countries")) +
    scale_x_log10() +
    theme(legend.position = "top") +
    guides(color = guide_legend(reverse = TRUE))
```

###Time for some new data

```{r}
organs.url <- "https://raw.githubusercontent.com/socviz/soc880/gh-pages/data/organ-procurement.csv"
organs.data <- read.csv(url(organs.url), row.names = 1)
```

Make sure R knows this is a date.

```{r}
organs.data$year <- as.Date(strptime(organs.data$year, format = "%Y"))
```

Playing around with the new data.

```{r}
new.plot <- ggplot(data = organs.data, aes(x = year, y = donors))
new.plot + geom_line(aes(group = country)) + facet_grid(~consent.law)

library(magrittr)
library(dplyr)

by.country <- organs.data %>% group_by(consent.law, country) %>%
              summarize(donors = mean(donors, na.rm = TRUE))
anew.plot <- ggplot(by.country, aes(x = donors, y = country, color = consent.law))
anew.plot + geom_point(size = 3)
```

That plot is out of order. Let's fix that. And make it pretty.

```{r}
anew.plot <- ggplot(by.country, aes(x = donors, y = reorder(country, donors)))
anew.plot + geom_point(size = 3) +
            facet_grid(consent.law ~ ., scales = "free") +
            labs(x = "Donor Procurement Rate (per million population)",
                 y = "") +
            theme(legend.position = "top")
```

Cleveland Dotplots are a good at plotting one-dimensional data with categories as an alternative to...

*Bar charts
*Column charts

Boxplots...

```{r}
third <- ggplot(data = organs.data, aes(x = reorder(country, donors, na.rm = TRUE), y = donors))
third + geom_boxplot(aes(fill = consent.law), outlier.color = "Black") +
        coord_flip() +
        labs(x = "", 
             y = "Donor Procurement Rate")
```

Maybe this would be better as a dotplot, since there's so few data points.

```{r}
dotp <- ggplot(data = organs.data, aes(x = reorder(country, donors, na.rm = TRUE), y = donors))

dotp + geom_jitter(aes(color = consent.law), position = position_jitter(width = 0.2)) + 
       coord_flip() +
       labs(x = "",
            y = "Donor Procurement Rate")
```